---
title: "QuantileMatrix"
author: "Daxing Xu"
date: "12/29/2020"
output: 
  html_document: # html_notebook, html_document
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: simplex # default, united, cerulean, journal, flatly, darkly, readable, spacelab, united, cosmo, lumen, paper, sandstone, simplex, and yeti
    highlight: haddock # default, tango, pygments, kate, monochrome, espresso, zenburn, haddock, and breezedark
    # bibliography: bibliography.json
---

```{r}
knitr::opts_chunk$set(eval = T,echo = F,include = T, warning = F,message = F)
# eval 代码要不要执行
# echo 代码要不要输出
# include 图要不要输出
# warning 警告要不要输出
# message 默认信息要不要输出(如bins=30)
library(tidyverse)
library(RColorBrewer)
library(ggpubr)
```

## Random select column

> -   zcat allIndividualTriadsBlockBySubspecies.txt.gz \| cut -f \$(shuf -i 5-6353 \| head -n 100 \| sort \| xargs echo \| tr [:blank:] ',') \> test.txt
> -   zcat allIndividualTriadsBlockBySubspecies.txt.gz \| cut -f 1,2,3,4 \> test0.txt
> -   paste test0.txt test.txt \> allIndividualTriadsBlockBySubspecies.sample100Column.txt

## Data import and mutate

```{r}
table <- read.delim("allIndividualTriadsBlockBySubspecies.sample100Column.txt")
infoDB <- read.delim("/Users/xudaxing/Documents/deleteriousMutation/002_vmapII_taxaGroup/taxa_InfoDB.txt")
infoDB <- infoDB %>% select(Taxa,fdBySubContinent) %>% rename(TaxonID=Taxa)
df <- table %>% 
  left_join(x=.,y=infoDB, by="TaxonID") %>%  mutate(Type=ifelse(is.na(fdBySubContinent),Group,fdBySubContinent)) %>%
  filter(Type %in% c("WE_AT","DE_AT","FT_AT","LR_EU","Cultivar")) %>%
  mutate(ifPseudo=ifelse(Group %in% c("WE_AT","DE_AT","FT_AT"),"pseudo","hexaploid")) %>%
  select(TaxonID,Group,AdditiveOrDominance,SlightlyOrStrongly,ifPseudo,fdBySubContinent,Type,5:105)
```

## Function

### Z-score function

```{r}
## add zscore column
## note: data.frame can only contain one triad
mutateZScore <- function(dataFrame, columnName){
  pseudoSummary_columnName <- dataFrame %>% 
  dplyr::group_by(ifPseudo, AdditiveOrDominance, SlightlyOrStrongly) %>% 
  summarise(pseudoMean=mean({{columnName}}),pseudoSD=sd({{columnName}})) %>% 
  filter(ifPseudo=="pseudo") %>% 
   ungroup() %>% 
  select(AdditiveOrDominance:pseudoSD)

  df_columnName <-  dataFrame %>% 
  # select(TaxonID:Type,{{columnName}}) %>% 
  left_join(y = pseudoSummary_columnName,by=c("AdditiveOrDominance", "SlightlyOrStrongly")) %>%   mutate(zscore=({{columnName}}-pseudoMean)/pseudoSD)
  df_columnName
}
```

### Quantile function

```{r}
## 根据ecdf函数和load值,  求出quantile
evaluateECDF <- function(ecdfColumn, value){
  res=vector(mode="numeric", length=length(value))
  for (i in seq_along(ecdfColumn)) {
    res[i]=ecdfColumn[[i]](value[i])
  }
  res
}

## note: data.frame can only contain one triad
mutateQuantile <- function(dataFrame, columnName){
  ecdf_columnName <- dataFrame %>% 
  filter(ifPseudo=="pseudo") %>% 
  dplyr::group_by(AdditiveOrDominance, SlightlyOrStrongly) %>%
  summarise(ecdfFun = list(ecdf=ecdf({{columnName}})))
  
  df_columnName <- dataFrame %>% 
  # select(TaxonID:Type,columnName) %>% 
  left_join(y = ecdf_columnName,by=c("AdditiveOrDominance", "SlightlyOrStrongly")) %>% 
  mutate(quantile=evaluateECDF(ecdfFun, {{columnName}}))
}
```

### T function

```{r}
## 根据t.test()计算t
calculatingT <- function(hexaploidLoad, pseudoLoad){
  res <- vector(mode = "numeric",length = length(hexaploidLoad))
  for (i in seq_along(hexaploidLoad)) {
    res[i] <- t.test(x = pseudoLoad[[i]], mu=hexaploidLoad[[i]])$statistic
  }
  res
}
calculatingP <- function(hexaploidLoad, pseudoLoad){
  res <- vector(mode = "numeric",length = length(hexaploidLoad))
  for (i in seq_along(hexaploidLoad)) {
    res[i] <- t.test(x = pseudoLoad[[i]], mu=hexaploidLoad[[i]])$p.value
  }
  res
}

mutateT <- function(dataFrame,columnName){
  pseudo <- dataFrame %>% 
    filter(ifPseudo=="pseudo") %>% 
    dplyr::group_by(AdditiveOrDominance, SlightlyOrStrongly) %>% 
    summarise(pseudo=list({{columnName}}))
  
  hexaploid <- dataFrame %>% 
    filter(ifPseudo=="hexaploid") %>% 
    dplyr::group_by(AdditiveOrDominance, SlightlyOrStrongly) %>%
    left_join(y = pseudo, by = c("AdditiveOrDominance","SlightlyOrStrongly")) %>% 
    dplyr::mutate(t=calculatingT({{columnName}}, pseudo), 
                  p=calculatingP({{columnName}}, pseudo))
  hexaploid
}
```

### Distribution of Load & z-score & quantile & t

```{r}
columns <- c("T000072","T001133","T003691","T005781")
 df_triad_t <- df %>% 
    select(TaxonID:Type,T000072) %>% 
    mutateZScore(T000072) %>% 
    mutateQuantile(T000072) %>% 
    mutateT(T000072) %>% 
    mutate(pvalue=ifelse(t>0,-p,p)) %>% 
    mutate(minusLogP=ifelse(t>0, log10(p), -log10(p))) %>% 
   select(TaxonID:Type,T000072,zscore,quantile,t,p,pvalue,minusLogP)
```


```{r}
columns <- c("T000072","T001133","T003691","T005781")
for (i in columns) {

  ## col: hexaploid and pseudo
  col <- c("red","#bdbdbd")
  
  ## extract triad df and mutate z-score & quantile
  df_triad <- df %>% 
  select(TaxonID:Type,all_of(i)) %>% 
  mutateZScore(.data[[i]]) %>% 
  mutateQuantile(.data[[i]])
  
  # loadZscoreQuantile 
  loadZscoreQuantile <- c(i, "zscore", "quantile")
  label <- c("Load","Z-score","Quantile")
  
  for (j in seq_along(loadZscoreQuantile)) {
    
    q <- ggplot(df_triad, aes(x=.data[[loadZscoreQuantile[j]]], 
                              fill=ifPseudo, color=ifPseudo))+
    geom_histogram(position = position_identity(), aes(y=..ndensity..),bins = 50,alpha=0.6)+
    facet_grid(AdditiveOrDominance~SlightlyOrStrongly)+
    scale_color_manual(values = col)+
    scale_fill_manual(values = col)+
    scale_y_continuous(breaks = seq(0,1,by=0.5))+
    scale_x_continuous()+
    # labs(x=label[j],title = i)+
    # labs(title = i)+
    # theme_minimal()+
    theme_classic()+
    theme(legend.position = "none", legend.title = element_blank(),
        text = element_text(size = 28),strip.background = element_blank(),
        axis.title.x = element_blank(),axis.line = element_blank(),
        panel.border = element_rect(colour = "black",fill = "transparent"))
    q
    # qRes[[count]] <- q
    outFile <- paste("figs/",i,".",label[j],".png",sep = "")
    ggsave(plot = q, filename = outFile, width = 8, height = 6,dpi = 300)
  }

  df_triad_t <- df %>% 
    select(TaxonID:Type,all_of(i)) %>% 
    mutateT(.data[[i]]) %>% 
    mutate(pvalue=ifelse(t>0,-p,p)) %>% 
    mutate(minusLogP=ifelse(t>0, log10(p), -log10(p))) %>% 
    select(TaxonID:Type,all_of(i),t,p,pvalue,minusLogP)
  
  tValue <- c("t","p","pvalue","minusLogP")
  
  for (j in seq_along(tValue)) {
    q <- ggplot(df_triad_t, aes(x=.data[[tValue[j]]],fill=ifPseudo, color=ifPseudo))+
    geom_histogram(position = position_identity(), aes(y=..ndensity..),bins = 50,alpha=0.6)+
    facet_grid(AdditiveOrDominance~SlightlyOrStrongly)+
    scale_color_manual(values = "blue")+
    scale_fill_manual(values = "blue")+
    scale_y_continuous(breaks = seq(0,1,by=0.5))+
    scale_x_continuous()+
    # labs(x="One sample T",title = i)+
    # labs(title = i)+
    # theme_minimal()+
    theme_classic()+
    theme(legend.position = "none", legend.title = element_blank(),
        text = element_text(size = 28),strip.background = element_blank(),
        axis.title.x = element_blank(),axis.line = element_blank(),
        panel.border = element_rect(colour = "black",fill = "transparent"))
    outFile <- paste("figs/",i,".",tValue[j],".oneSampleT",".png",sep = "")
    ggsave(plot = q, filename = outFile, width = 8, height = 6,dpi = 300)
  }
}
```
